---
title: "2022 Spring Project"
author: "Justin Ehly"
date: "3/31/2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Project Scope:
* Use the train climate series data from New Delhi 2013 - 2017 to predict the temperature
* Data is pre-separated into Train and Test
* Source: https://www.kaggle.com/datasets/sumanthvrao/train-climate-Days-series-data?select=trainDelhiClimateTrain.csv

#######

import training/ testing data

```{r}
#setwd('C:/Users/justi/Google Drive/_SMU/6373 Days Series/2022 Project')

train = read.csv('DailyDelhiClimateTrain.csv') # training data
test = read.csv(('DailyDelhiClimateTest.csv')) # testing data
head(train[-1])
tail(train)
train = head(train,-1)  # last row seems to be mistake
tail(train)
head(test)


train = rbind(train, test) # merge dataframes and re-split them

# note, this didn't end up getting used, but it was going to be too much work to go back and fix it
test = tail(train, 21)  
train = head(train, - 21) # remove test data


summary(train)

library(naniar)
library(tswge)

##### Look for any missing values including NA, variants of NA and 'UNKNOWN's in the data set ####
missing <- sapply(train, function(x) sum(sum(x %in% common_na_strings), 
                                        sum(x %in% common_na_numbers),
                                        sum(is.na(x)), 
                                        sum(x %in% 'UNKNOWN')))
missing # shows 13 missing from humidity
#common_na_numbers # -9   -99  -999 -9999  9999    66    77    88
#train$humidity
# very possible the common_na_numbers are picking up humidity readings that are actual and not missing
```

plot data
```{r plot the data}
par(mar=c(1,1,1,1))
par(mfrow = c(4,1))
plot(train$meantemp, type='l', main = 'Mean Temp', col = 'black')
plot(train$humidity, type='l', main = 'Humidity', col = 'blue')
plot(train$wind_speed, type = 'l', main = 'Wind Speed', col = 'green')
plot(train$meanpressure, type = 'l', main = 'Mean Pressure', col = 'red')

library(GGally)
ggpairs(train[,2:5]) + theme_bw()

```

visually inspect for stationarity for each element

```{r plot mean temp}
# temp
dev.off()
Days = seq(1,1554)
plot(Days, train$meantemp, type='l', lwd=1, col='darkorange3', pch=20, ylab='Temperature')
title(main='2013-2017 Daily Temperature in New Delhi', col.main = 'darkorange3')

acf(train$meantemp, col='darkorange3', lwd = 3, main='')
title(main='ACF 2013-2017 Daily Temperature in New Delhi', col.main = 'darkorange3')

pt = plotts.sample.wge(train$meantemp)
dev.off()
parz = parzen.wge(train$meantemp, trunc=365, plot=F)
plot(parz$freq, parz$pzgram, xlab='Frequency', ylab='dB', col = 'darkorange3', type='l')
title(main='Parzen Window Tuncation Point: M = 365', col.main = 'darkorange3')

per = period.wge(train$meantemp)
per$pgram[1:10]  # strongest  = 27.xxx in index 4
per$freq[4] # 0.002574003

```

Mean Temperature does not appear to be stationary due to it’s slowly damping ACF…it is temperature data and from experience we know it is seasonal. We also see a peak at about 0.83 (12 months) and then additional peaks.
We also see sinusoidal pattern in the realization

```{r plot humidity}
# humidity
ph = plotts.sample.wge(train$humidity)

```

Humidity similarly does not appear to be stationary due to it’s sinusoidal pattern in the realization and slowly damping ACF…it is temperature data and from experience we know it is seasonal. We also see a peak at about 0.83 (12 months) and then additional peaks


```{r plot wind speed}
# wind speed
pws = plotts.sample.wge(train$wind_speed)

```

train Wind Speed also does not appear to be stationary due to the cyclical pattern in the realization, the slowly damping, but almost sinusoidal pattern in the acf and the peaks at 0, 0.83 and etc in the spectral density

```{r plot mean pressure}
# mean pressure
pmp = plotts.sample.wge(train$meanpressure, arlimits = T)


```
Mean train pressure does appear to be stationary with the randomness in the realization, the quickly damping to near zero in the afc.



```{r Differencing Data}
# temp
# differces of 1, 12
temp_diff1 = artrans.wge(train$meantemp, phi.tr = c(1), plottr = F)
dev.off()
Time = seq(1,1553)
par(mfrow = c(1,2))
plot(Time,temp_diff1, col = 'darkorange3', type='l', ylab='Temperature')
title(main='Realization 1st Difference', col.main = 'darkorange3')
acf(temp_diff1, col='darkorange3', lwd = 3, main='')
title(main='ACF 1st Difference', col.main = 'darkorange3')

dev.off()
temp_365 = artrans.wge(train$meantemp, phi.tr = 365, plottr = F)
acf(temp_365, col='darkorange3', lwd = 3, main='')
title(main='ACF 365th Difference', col.main = 'darkorange3')


temp_1_365 = artrans.wge(temp_diff1, phi.tr = 365, plottr = F)
acf(temp_1_365)

temp_diff1_5 = artrans.wge(temp_diff1, phi.tr = c(rep(0,4),1))
acf(temp_diff1_5)

temp_388_5 = artrans.wge(train$meantemp, phi.tr = 388.5)

```
Looks like we potentially have a (1-B) in the model
With additional tests, we can confirm using the ACFs that there is no differencing at 3,4,6, 12, 30 or 90
```{r}

plotts.sample.wge(temp_diff1)

```
Interesting, with a single difference, we see a big dip at about .012, suggesting there is an MA component in this data after a difference.





### ARMA Model ###

```{r ARMA}
aic5.wge(train$meantemp, p=0:8, q=0:3) #picks 6,2
fit_arma = est.arma.wge(train$meantemp, p = 6, q = 2)
mean(train$meantemp)
fit_arma$avar
```
Notice the first factor is nearly on the unit circle, again I think this should be differenced at least once for any sort of good forecast.

```{r}
# generate and plot 3 sets of data using the nonstationary model


g = list()
for (i in 1:3){
  g[[i]] = gen.arima.wge(200, phi = fit_arma$phi, theta = fit_arma$theta, vara = fit_arma$avar)
  g[[i]] = g[[i]] + mean(train$meantemp)
}
dev.off()
par(mfrow = c(4,1))
plot(train$meantemp[1:200], type = 'l', main = 'Original Temp Data')
plot(g[[1]], type = 'l', main = 'Generated Data')
plot(g[[2]], type = 'l', main = 'Generated Data')
plot(g[[3]], type = 'l', main = 'Generated Data')

```

```{r train test split}

# hold back 200 days for cross validation
x_test = tail(train[-1],200)

X = cbind(head(train[-1],-200))
```


```{r 21-day forecast ARMA62}
# 7 - day forecast
fore_arma = fore.arma.wge(X$meantemp, phi = fit_arma$phi, theta =  fit_arma$theta, 
                          n.ahead = 21, lastn = FALSE, plot=FALSE)
```

```{r plot 21day ARMA(6,2)}
Days = seq(1255,1375)
dev.off()
plot(Days, train$meantemp[1255:1375], type = 'b', ylim = c(20,37), col='darkgrey', main = 'ARMA(6,2) 21-Day Temp Forecast', ylab = 'Temperature')
lines(Days[101:121], fore_arma$f, type = 'l',  lwd = 5, col = 'red', pch=16 )
lines(Days[101:121], fore_arma$ll, type = 'l',  lwd = 3,  col = 'blue', pch=16 )
lines(Days[101:121], fore_arma$ul, type = 'l',  lwd = 2,  col = 'blue', pch=16 )
```
```{r ARMA(6,2) performance}
arma_ase = mean((x_test$meantemp[1:21] - fore_arma$f)^2)
arma_ase
arma_rmse = sqrt(arma_ase)
arma_rmse
```
Not a bad 7-day, but definitely lower than the actual data with the upper limit suggesting that we are within range of the original data.
ASE: 9.123
RMSE: 3.02


```{r resids white noise}
# check residuals for white noise
parma = plotts.sample.wge(fore_arma$resid, arlimits = T)
ljung.wge(fore_arma$resid, p = 7, q = 2, K = 24)
ljung.wge(fore_arma$resid, p = 7, q = 2, K = 48)
```
Test for white noise from residuals: 
Visually looks like white noise
Ljung test suggest this is white noise with pvalues at .25 and .28 for k=24 and 48 respectively.

```{r 200-day foreast ARMA62}
# 21 - day forecast
fore_arma_200 = fore.arma.wge(X$meantemp, phi = fit_arma$phi, theta =  fit_arma$theta, n.ahead = 200, lastn =FALSE, plot=FALSE)
```

```{r plot 200}
Days = seq(1:1554)
# plot the forecast 
dev.off()
plot(Days, train$meantemp[1:1554], type = 'l', main = 'ARMA(6,2) 200-Day Temp Forecast', ylab = 'Temperture', ylim = c(0,40), col = 'darkgrey')
lines(Days[1355:1554], fore_arma_200$f, col = 'red', type='l', lwd = 5)
lines(Days[1355:1554], fore_arma_200$ll, col = 'blue', type = 'l', lwd = 5)
lines(Days[1355:1554], fore_arma_200$ul, col = 'blue', type = 'l', lwd = 5)

```


```{r stats}
arma_200_ase = mean((x_test$meantemp - fore_arma_200$f)^2)
arma_200_rmse = sqrt(arma_200_ase)
arma_200_ase
arma_200_rmse

```
Forecast actually doesn't look too bad for a 200-day forecast! It follows the curvature of the data, but doesn't necessarily pick up on the smaller variance.
ASE: 9.882!, RMSE: 3.143
actually might be a really good model!



```{r rolling window RMSE}
# rolling window forecasts

roll_arma_21_rmse = roll.win.rmse.wge(train$meantemp, horizon = 21, phi = fit_arma$phi, theta = fit_arma$theta)
roll_arma_200_rmse = roll.win.rmse.wge(train$meantemp, horizon = 200, phi = fit_arma$phi, theta = fit_arma$theta)



```
21-day Rolling RMSE: 137.765
200-day Rolling RMSE: 298.491







### VAR Method ###

VAR Method
cross correlation function
```{r}
dev.off()
ccf(train$meanpressure, train$meantemp) # because i suspect there is a lagged relationship b/w pressure and temp
```
No significant lags here between pressure and temperature, although I thought there might be

```{r}
dev.off()
ccf(train$humidity, train$meantemp)
```
Main spike is at 0, suggesting that humidity and temp coexist

```{r}
dev.off()
ccf(train$wind_speed, train$meantemp)
ccf(train$wind_speed, train$meantemp, lag.max = 100)
```
This sounds crazy, but after a little reading on google about wind, pressure and temperatue it is possible there should be a lag on wind at -28. But a little curiousity and we set the lag.max to 100 and we see this might not be useful information.


```{r}
library(vars)
# hold back 200 days for cross validation
x_test = tail(train[-1],200)

X = cbind(head(train[-1],-200))
vs = VARselect(X, lag.max = 15, type = 'const', season=NULL, exogen = NULL) # AIC 6, HQ(n) 4, BIC 1, FEP(n) 6
vs$selection  # selection
vs$criteria

vs_sea = VARselect(X, lag.max = 15, type = 'const', season = 365) # AIC 5, BIC 1
vs_sea$criteria
vs_sea$selection

min(vs$criteria[1,])  #mean AIC
# VARsselect picks p=6 mean AIC 17.543
lsfit6 = VAR(X, p=6, type = 'const')
lsfit4 = VAR(X, p=4, type = 'const')
lsfit1 = VAR(X, p=1, type = 'const')
lsfit_sea =  VAR(X, p=5, type = 'const', season = 365)

```

From the equation representation we can see that 
meantemp.l2, meantemp.l3 and  wind_speed.l6  are the only significant relationships to temp at X_t1

```{r forecast 21-day}
preds6 = predict(lsfit6, n.ahead = 21)
preds4 = predict(lsfit4, n.ahead = 21)
preds1 = predict(lsfit1, n.ahead = 21)
preds_sea = predict(lsfit_sea, n.ahead = 21)
mean(X$meantemp)
```

```{r 21-day permformance}

var6_21_ase = mean((x_test$meantemp[1:21] - preds6$fcst$meantemp[,1])^2)
var4_21_ase = mean((x_test$meantemp[1:21] - preds4$fcst$meantemp[,1])^2)
var1_21_ase = mean((x_test$meantemp[1:21] - preds1$fcst$meantemp[,1])^2)
var_sea_21_ase = mean((x_test$meantemp[1:21] - preds_sea$fcst$meantemp[,1])^2)

var6_21_rmse = sqrt(var6_21_ase)
var4_21_rmse = sqrt(var4_21_ase)
var1_21_rmse = sqrt(var1_21_ase)
var_sea_21_rmse = sqrt(var_sea_21_ase)

var6_21_ase
var4_21_ase
var1_21_ase
var_sea_21_ase

var6_21_rmse
var4_21_rmse
var1_21_rmse
var_sea_21_rmse
```
VAR6/ ASE:1.46 RMSE: 1.208
VAR4/ ASE:1.724 RMSE: 1.313
VAR1/ ASE:1.46 RMSE: 1.208
VAR Seasonal/ ASE:7.24 RMSE: 2.692


```{r plot}
dev.off()
Days = seq(1255,1375)
plot(Days, train$meantemp[1255:1375], type = 'b', ylim = c(27,37), col='darkgrey',
     main = 'VAR 21-Day Temp Forecast', ylab = 'Temperature')
lines(Days[101:121], preds6$fcst$meantemp[,1], type = 'l',  lwd = 5, col = 'red', pch=16 )
lines(Days[101:121], preds4$fcst$meantemp[,1], type = 'l',  lwd = 3,  col = 'green', pch=16 )
lines(Days[101:121], preds1$fcst$meantemp[,1], type = 'l',  lwd = 2,  col = 'blue', pch=16 )
lines(Days[101:121], preds_sea$fcst$meantemp[,1], type = 'l',  lwd = 2,  col = 'purple', pch=16 )
legend(x = 1349, y = 37, legend = c('VAR P=6','VAR P=4', 'VAR P=1', 'VAR S=365 P=5'),
       col = c('red', 'green', 'blue', 'purple'), lty = 1, lwd = 3,cex = 0.8)
```
VAR without seasonality doesn't do much to reflect the variability in the data. With the seasonality at 365, we see the variability accounted for, but the forecasts seem to be closer to mu = 25.59. The non seasonal forecasts appear to be damping towards the mean.

```{r plot VAR1 alone}
Days = seq(1255,1375)
dev.off()

plot(Days, train$meantemp[1255:1375], type = 'b', ylim = c(21,40), col='darkgrey', main = 'VAR1 21-Day Temp Forecast', ylab = 'Temperature')
lines(Days[101:121], preds1$fcst$meantemp[,1], type = 'l',  lwd = 2,  col = 'red', pch=16 )
lines(Days[101:121], preds1$fcst$meantemp[,2], type = 'l',  lwd = 2,  col = 'blue', pch=16 )
lines(Days[101:121], preds1$fcst$meantemp[,3], type = 'l',  lwd = 2,  col = 'blue', pch=16 )
```

```{r 200-day forecast}

preds2006 = predict(lsfit6, n.ahead = 200)
preds2004 = predict(lsfit4, n.ahead = 200)
preds2001 = predict(lsfit1, n.ahead = 200)
preds200_sea = predict(lsfit_sea, n.ahead = 200)
```

```{r plot 200 fir var}
Days = seq(1,1554)
dev.off()
plot(Days, train$meantemp, type = 'l', ylim = c(7,40), col='darkgrey', main = 'VAR 200-Day Temp Forecast', ylab = 'Temperature')
lines(Days[1355:1554], preds2006$fcst$meantemp[,1], type = 'l', lwd = 7, col = 'red', pch=16 )
lines(Days[1355:1554], preds2004$fcst$meantemp[,1], type = 'l', lwd = 5,  col = 'green', pch=16 )
lines(Days[1355:1554], preds2001$fcst$meantemp[,1], type = 'l', lwd = 2,  col = 'blue', pch=16 )
lines(Days[1355:1554], preds200_sea$fcst$meantemp[,1], type = 'l', lwd = 2,  col = 'purple', pch=16 )
legend(x = 1210, y = 40, legend = c('VAR P=6','VAR P=4', 'VAR P=1', 'VAR S = 365 P = 5'),
       col = c('red', 'green', 'blue', 'purple'), lty = 1, lwd = 3,cex = 0.8)
```
Seasonal VAR seems to have NAILED it!

```{r plot seasonal model alone}
Days = seq(1,1554)
dev.off()
plot(Days, train$meantemp, type = 'l', ylim = c(7,40), col='darkgrey', main = 'VAR P=5, S=365 200-Day Temp Forecast', ylab = 'Temperature')
lines(Days[1355:1554], preds200_sea$fcst$meantemp[,1], type = 'l', lwd = 2, col = 'red', pch=16 )
lines(Days[1355:1554], preds200_sea$fcst$meantemp[,2], type = 'l', lwd = 2,  col = 'blue', pch=16 )
lines(Days[1355:1554], preds200_sea$fcst$meantemp[,3], type = 'l', lwd = 2,  col = 'blue', pch=16 )


```


```{r}
fanchart(preds2001, colors =c('darkorange3'), col.y = 'darkgrey')

```

```{r}
var2006_ase = mean((x_test$meantemp - preds2006$fcst$meantemp[,1])^2)
var2004_ase = mean((x_test$meantemp - preds2004$fcst$meantemp[,1])^2)
var2001_ase = mean((x_test$meantemp - preds2001$fcst$meantemp[,1])^2)
var200_sea_ase = mean((x_test$meantemp - preds200_sea$fcst$meantemp[,1])^2)
var2006_ase
var2004_ase
var2001_ase
var200_sea_ase
var2006_rmse = sqrt(mean((x_test$meantemp - preds2006$fcst$meantemp[,1])^2))
var2004_rmse = sqrt(mean((x_test$meantemp - preds2004$fcst$meantemp[,1])^2))
var2001_rmse = sqrt(mean((x_test$meantemp - preds2001$fcst$meantemp[,1])^2))
var200_sea_RMSE = sqrt(var200_sea_ase)
var2006_rmse
var2004_rmse
var2001_rmse
var200_sea_RMSE

```
p=6/ ASE: 73.06, RMSE: 8.55 (AIC)
p=4/ ASE: 68.77, RMSE: 8.29 (HQ(n))
p=1/ ASE: 52.12, RMSE: 7.22 (BIC)
S=365, p=5/ ASE: 9.39, RMSE: 3.06



```{r}
# white noise test 21-day VAR(p=1)
pvar = plotts.sample.wge(lsfit1$varresult$meantemp$residuals, arlimits = T)
ljung.wge(lsfit1$varresult$meantemp$residuals, p=1, K=24)
ljung.wge(lsfit1$varresult$meantemp$residuals, p=1, K=48)
```
VAR with p=6
The residuals do not appear to be white noise and the pvalues of the ljung test confirm they are not with pvalues = 0.032 and 0.044 at K=24 and 48

```{r}
# white noise test for the best performing VAR model
pvar = plotts.sample.wge(lsfit_sea$varresult$meantemp$residuals, arlimits = T)
ljung.wge(lsfit_sea$varresult$meantemp$residuals, p=5, K=24)
ljung.wge(lsfit_sea$varresult$meantemp$residuals, p=5, K=48)
```
VAR with S=365 and p=5
Visually the residuals appear to be white noise.
Using the Ljung Box test, we confirm that residuals are white noise with pvalues of 0.395 and 0.066 with K = 24 and 48 respectively.
This model is USEFUL!

###########

### MLP ###


```{r base fit}

library(nnfor)
# train/ test split
x_test = tail(train[-1],200)
X = cbind(head(train[-1],-200))

xrg = data.frame(humity = ts(X$humidity), wind_speed = ts(X$wind_speed), meanpressure =  ts(X$meanpressure))

basefit = mlp(ts(X$meantemp), xreg = xrg, comb = 'median')
plot(basefit)
basefit
```

```{r custom fit}
mfit = mlp(ts(X$meantemp), lags = c(1,2,3,4,5,6,7), xreg = xrg, comb='median')
plot(mfit)
mfit
```


```{r seasonal fit}
seasonalfit = mlp(ts(X$meantemp), lags = c(1,2,3,4,5,6,7), xreg = xrg, comb='median', 
                  allow.det.season = TRUE)
plot(seasonalfit)
seasonalfit
```


```{r}
# forecast each model 21-days
# create xreg data frame by adding 21 values from the beginning of the test set of data to the training set
xrg_fore =  data.frame(humity = ts(c(X$humidity, head(x_test$humidity,21))), 
                       wind_speed = ts(c(X$wind_speed, head(x_test$wind_speed,21))), 
                       meanpressure =  ts(c(X$meanpressure,head(x_test$meanpressure,21))))

fore_basefit = forecast(basefit, h = 21, xreg = xrg_fore)
plot(fore_basefit)
```

```{r}
fore_mfit = forecast(mfit, h = 21, xreg = xrg_fore)
plot(fore_mfit)

```

```{r}
fore_seasonalfit = forecast(seasonalfit, h = 21, xreg = xrg_fore, level = 95)
plot(fore_seasonalfit)
```

```{r 21-day mlp pfrm}
mlpbase_21_ase = mean((x_test$meantemp[1:21] - fore_basefit$mean)^2)
mlpmfit_21_ase = mean((x_test$meantemp[1:21] - fore_mfit$mean)^2)
mlpseasonal_21_ase = mean((x_test$meantemp[1:21] - fore_seasonalfit$mean)^2)

mlpbase_21_rmse = sqrt(mlpbase_21_ase)
mlpmfit_21_rmse = sqrt(mlpmfit_21_ase)
mlpseasonal_21_rmse = sqrt(mlpseasonal_21_ase)

mlpbase_21_ase
mlpmfit_21_ase
mlpseasonal_21_ase
mlpbase_21_rmse
mlpmfit_21_rmse
mlpseasonal_21_rmse

```

MLP BASE ASE: 2.017, RMSE: 1.42
MLP FIT ASE: 2.741, RMSE: 1.656
MLP Seasonal ASE: 1.957, RMSE: 1.399


```{r mlp 21-day forecast plot}


Days = seq(1255,1375)
dev.off()
plot(Days, train$meantemp[1255:1375], type = 'l', ylim = c(27,37), col='brown', 
     main = 'MLP 21-Day Temp Forecast', ylab = 'Temperature')
lines(Days[100:120], fore_basefit$mean, type = 'l', lwd = 3, col = 'red', pch=16 )
lines(Days[100:120], fore_mfit$mean, type = 'l', lwd = 3,col = 'green', pch=16 )
lines(Days[100:120], fore_seasonalfit$mean, type = 'l', lwd = 3, col = 'blue', pch=16 )
legend(x = 1349, y = 37, legend = c('Base MLP','Custom Tuned MLP', 'Seasonal MLP'),
       col = c('red', 'green', 'blue'), lty = 1, lwd = 3,cex = 0.8)
```


# forecast each model 200-days
```{r}
# create xreg data frame by adding 21 values from the beginning of the test set of data to the training set

xrg_fore =  data.frame(humity = ts(train$humidity), 
                       wind_speed = ts(train$wind_speed), 
                       meanpressure =  ts(train$meanpressure))

fore_basefit200 = forecast(basefit, h = 200, xreg = xrg_fore)
plot(fore_basefit200)
```

```{r}
fore_mfit200 = forecast(mfit, h = 200, xreg = xrg_fore)
plot(fore_mfit200)

```

```{r}
fore_seasonalfit200 = forecast(seasonalfit, h = 200, xreg = xrg_fore)
plot(fore_seasonalfit200)

```


```{r}

Days = seq(1000,1554)
dev.off()
plot(Days, train$meantemp[1000:1554], type = 'l', ylim = c(7,40), col='brown', main = 'MLP 200-Day Temp Forecast', ylab = 'Temperature')
lines(Days[355:554], fore_basefit200$mean, type = 'l', lwd = 3, col = 'red', pch=16 )
lines(Days[355:554], fore_mfit200$mean, type = 'l', lwd = 3,col = 'green', pch=16 )
lines(Days[355:554], fore_seasonalfit200$mean, type = 'l', lwd = 3, col = 'blue', pch=16 )
legend(x = 1430, y = 40, legend = c('Base MLP','Custom Tuned MLP', 'Seasonal MLP'),
       col = c('red', 'green', 'blue'), lty = 1, lwd = 3,cex = 0.8)
```


```{r}

mlpbase_200_ase = mean((x_test$meantemp - fore_basefit200$mean)^2)
mlpmfit_200_ase = mean((x_test$meantemp - fore_mfit200$mean)^2)
mlpseasonal_200_ase = mean((x_test$meantemp - fore_seasonalfit200$mean)^2)


mlpbase_200_rmse = sqrt(mean((x_test$meantemp - fore_basefit200$mean)^2))
mlpmfit_200_rmse = sqrt(mean((x_test$meantemp - fore_mfit200$mean)^2))
mlpseasonal_200_rmse = sqrt(mean((x_test$meantemp - fore_seasonalfit200$mean)^2))

mlpbase_200_ase
mlpmfit_200_ase
mlpseasonal_200_ase

mlpbase_200_rmse
mlpmfit_200_rmse
mlpseasonal_200_rmse

```
base mlp/ ASE: 104.36, RMSE: 10.21
mfit mlp/ 102.96, RMSE: 10.03
seasonal mlp/ ASE: 102.48 RMSE: 10.15

```{r ljung for nn}
# Bivin I borrowed your code

ljung.nn.wge=function (x, K=24, hd=0) 
{
    cor <- acf(x, lag.max = K, plot = FALSE, na.action = na.pass)
    n <- length(x)
    df=K-hd
    obs <- cor$acf[2:(K + 1)]
cat('Obs',obs,'\n')
    test <- "Ljung-Box test"
    chi.square <- n * (n + 2) * sum(1/seq.int(n - 1, n - K) * obs^2)
    pval <- 1 - pchisq(chi.square, K - hd)
out1=list(test=test,K=K, chi.square=chi.square,df=df,pval=pval)
return(out1)   
}
```

```{r}
# check for white noise
mlpp = plotts.sample.wge(fore_seasonalfit200$residuals, arlimits = T)
ljung.nn.wge(fore_seasonalfit200$residuals, K = 24, hd = seasonalfit$hd)
ljung.nn.wge(fore_seasonalfit200$residuals, K = 48, hd = seasonalfit$hd)

```
The seasonal fit residuals for the mlp visually appear to be white noise.
If I correctly adapted the ljung function for the mlp, using the number of hidden nodes in place of the p and q from the ARMA models, then we can confirm the residuals are white noise with pvalues pf 0.0586 and 0.142 with K=24 and K=48 respectively.

### Ensemble ###

```{r ensemble}
ensemble_21day = list()
ensemble_200day = list()
ensemble_21day$f = (fore_arma$f + preds_sea$fcst$meantemp[,1])/2
ensemble_21day$ul = (fore_arma$ul + preds_sea$fcst$meantemp[,3])/2
ensemble_21day$ll = (fore_arma$ll + preds_sea$fcst$meantemp[,2])/2

ensemble_200day$f = (fore_arma_200$f +  preds200_sea$fcst$meantemp[,1]) /2
ensemble_200day$ul = (fore_arma_200$ul +  preds200_sea$fcst$meantemp[,3]) /2
ensemble_200day$ll = (fore_arma_200$ll +  preds200_sea$fcst$meantemp[,2]) /2
```


```{r mlp 21-day forecast plots}


Days = seq(1255,1375)
dev.off()
plot(Days, train$meantemp[1255:1375], type = 'l', ylim = c(20,37), col='brown', 
     main = 'Ensemble (ARMA(6,2) + VAR(S=365,p=5)) Model 21-Day Temp Forecast', ylab = 'Temperature')
lines(Days[100:120], ensemble_21day$f, type = 'l', lwd = 3, col = 'red', pch=16 )
lines(Days[100:120], ensemble_21day$ul, type = 'l', lwd = 3,col = 'blue', pch=16 )
lines(Days[100:120], ensemble_21day$ll, type = 'l', lwd = 3, col = 'blue', pch=16 )

```

```{r ensemble permormance}

ens_21_ase = mean((x_test$meantemp[1:21] - ensemble_21day$f)^2)
ens_21_rmse = sqrt(ens_21_ase)
ens_21_ase
ens_21_rmse

```
Enemble 21-d ASE: 7.992
Ensemble 200-d RMSE: 1.975

```{r ensemble 200-day plot}

Days = seq(1,1554)
dev.off()
plot(Days, train$meantemp, type = 'l', ylim = c(0,40), col='darkgrey', 
     main = 'Ensemble (ARMA(6,2) + VAR(S=365, p=5) Model 200-Day Temp Forecast', ylab = 'Temperature')
lines(Days[1355:1554], ensemble_200day$f, type = 'l', lwd = 3, col = 'red', pch=16 )
lines(Days[1355:1554], ensemble_200day$ul, type = 'l', lwd = 3,col = 'blue', pch=16 )
lines(Days[1355:1554], ensemble_200day$ll, type = 'l', lwd = 3, col = 'blue', pch=16 )
```

```{r 200day perfm}

ens_200_ase = mean((x_test$meantemp - ensemble_200day$f)^2)
ens_200_rmse = sqrt(ens_200_ase)
ens_200_ase
ens_200_rmse
```
Ensemble 200-d ASE: 6.448
Ensemble 200-d RMSE: 1.975

### Final Model Stats ###
```{r final stats}
model_performance = data.frame(Models = c('ARMA(6,2)','VAR(p=1)', 'VAR(S=365,p=5)', 'MLP-Seasonal', 'Ensemble'),
                               'ASE 21-Day' = c(arma_ase, var1_21_ase, var_sea_21_ase, mlpseasonal_21_ase, ens_21_ase),
                               'RMSE 21-Day' = c(arma_rmse, var1_21_rmse, var_sea_21_rmse, mlpseasonal_21_rmse, 
                                                 ens_21_rmse),
                               'ASE 200-Day' = c(arma_200_ase, var2001_ase, var200_sea_ase, mlpseasonal_200_ase,
                                                 ens_200_ase),
                               'RMSE 200-Day' = c(arma_200_rmse, var2001_rmse, var200_sea_RMSE, mlpseasonal_200_rmse,
                                                  ens_200_rmse))

```

```{r}

model_performance

```


